---
title: "ScRNAseq analysis [Library 3]{style='color:red ; font-weight:500'}"
subtitle: "[Processing]{style='color:yellow ; font-weight:600 ; background-color:black ; font-size:120%'}"
author: "juba"
date: "`r Sys.Date()`"
output: 
  html_document:
    number_sections: no
    theme: flatly
    highlight: tango
    toc: yes
    toc_depth: 6
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged 

params:
  c3m: TRUE
---

```{r Setting, echo=FALSE, message=FALSE, warning=FALSE}

##########################
#####   Packages    ######
##########################

packages <- installed.packages()[,"Package"]
cran_packages <- c("knitr", "Seurat", "dplyr", "ggplot2", "tibble", 
                   "writexl","clustree","cowplot") 
bioc_packages <- c("clusterProfiler", "org.Hs.eg.db", "AnnotationDbi", 
                   "fgsea", "enrichplot","ComplexHeatmap")

## Installing cran packages : 
for(package in cran_packages[!(cran_packages %in% packages)]) {
  install.packages(package)
}

## Install bioconductor packages : 
for(package in bioc_packages[!(bioc_packages %in% packages)]) {
  BiocManager::install(package)
}

## Loading packages
library(knitr)
library(Seurat)
library(dplyr)
library(ggplot2)
library(tibble)
library(writexl)
library(clustree)
library(stringr)
library(cowplot)
library(devtools)



##################################
#######   Other settings  ########
##################################

## Palettes :
MyPalette <- c("#9933aa","#ffdd22","#aa4400","#ff0000","#337722","#00ff66","#005566","#002277",
               "#441144","#aa0077","#00bbff","#003333","#4422cc","#116611","#330077","#111111",
               "#667700","#ddaa00","#33ffff","#ff22ff","#ffff33","#00ff00","#0000ff","#444444")

## Directories :
if(params$c3m == T){
  out_object_dir <- "~/Bureau/Projects/SingleCell/ScAnalysis/Library3/Out_objects/"
  out_figure_dir <- "~/Bureau/Projects/SingleCell/ScAnalysis/Library3/Out_figures/"
  out_seurat_dir <- "~/Bureau/Projects/SingleCell/ScAnalysis/Library3/SeuratObjects/"
} else {
  out_object_dir <- "/Projects/T-ALL scRNAseq/Library3/OutObjects/"
  out_figure_dir <- "/Projects/T-ALL scRNAseq/Library3/OutFigures/"
  out_seurat_dir <- "/Projects/T-ALL scRNAseq/Library3/SeuratObjects/"
}


## Set chunk options
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
```

```{r}
Assign_colors <- function(object, pal){
  n = length(unique(object))
  
  if(is.numeric(object)){
    message("This function can't handle numeric argument !!!")
    stop("Please provide a character vector as an object")
  }
  
  if(length(pal) < n){
    warning("Not enough colours within the selected palette to match your variables !!")
    warning("Generate colours from the implemented list within base R")
    pal <- sample(colours(), size = n)
  }
  
  res = sample(pal, size = n) %>% setNames(unique(object))
  
  return(res)
}

My_ggplot_theme <- function(){
  theme_linedraw()+
    theme(panel.background = element_rect(fill = "#fff", colour = "#aaa"),
        panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.title.x = element_text(size = 14, margin = margin(t=0.15, unit = "in")),
        axis.title.y = element_text(size = 14, margin = margin(r=0.15, unit = "in")),
        legend.key = element_blank(),
        legend.background = element_blank(),
        legend.title = element_text(size = 13, face = "bold", colour = "darkred", 
                                    margin = margin(b=0.15, unit = "in")),
        legend.text = element_text(size = 11),
        legend.box.margin = margin(l= 0.05, unit = "in"),
        plot.title = element_text(size = 15, face = "bold", colour = "darkred"))
}
```


#   Import the data   {.tabset .tabset-pills}

```{r}
if(params$c3m == T){
  SeuratObj <- SeuratObj <- readRDS("~/Bureau/Projects/SingleCell/ScAnalysis/Library3/SeuratObjects/Seurat2.rds")
} else {
  SeuratObj <- readRDS("/Projects/T-ALL scRNAseq/Library3/SeuratObjects/ Seurat2.rds")
}

SeuratObj
```

```{r ,echo=FALSE}
SeuratObj@meta.data <- SeuratObj@meta.data %>% 
  mutate(hash.ID = case_when(
    hash.ID == "M187" ~ "Diagnosis",
    hash.ID == "M187r" ~ "Relapse"
  ))

SeuratObj@meta.data <- SeuratObj@meta.data %>% 
  mutate(
    Clust_res0.1 = as.factor(paste0(SCT_snn_res.0.1, substr(SeuratObj$hash.ID,1,1))),
    Clust_res0.3 = as.factor(paste0(SCT_snn_res.0.3, substr(SeuratObj$hash.ID,1,1))),
    Clust_res0.5 = as.factor(paste0(SCT_snn_res.0.5, substr(SeuratObj$hash.ID,1,1))),
    Clust_res0.7 = as.factor(paste0(SCT_snn_res.0.7, substr(SeuratObj$hash.ID,1,1))),
    Clust_res0.9 = as.factor(paste0(SCT_snn_res.0.9, substr(SeuratObj$hash.ID,1,1))),
    Clust_res1.1 = as.factor(paste0(SCT_snn_res.1.1, substr(SeuratObj$hash.ID,1,1)))
  )
```


[Here're some representation]{style='color:darkred ; font-weight:600 ; font-size:120%'}

##    General conditions

```{r}
DimPlot(SeuratObj, cols = MyPalette, group.by = "hash.ID", label = F, label.box = F, repel = F)
```


##    Clusteres at res0.3

```{r}
DimPlot(SeuratObj, cols = MyPalette, group.by = "Clust_res0.3", 
        label = T, label.box = T, repel = T, label.color = "white", label.size = 4)
```

##    Clonotypes distributions

```{r}
DimPlot(SeuratObj, group.by = "clone_id")+
  theme(legend.position = "none")
```


##    Cell cycle distribution

```{r}
DimPlot(SeuratObj, cols = MyPalette, group.by = "Phase", 
        label = T, label.box = T, repel = T, label.color = "white", label.size = 4)
```




#   Cells inspection    {.tabset .tabset-pills}

[I'll inspect the expression of some markers to determine the cell types]{style='color:darkred ; font-weight:600 ; font-size:120%'}

This will help us determine any cluster that may not be of interest !

[We'll realize the inspection using the expression plot and violin plots with the 0.2 resolution below]{style='color:purple ; font-weight:600 ; font-size:100%'}

```{r}
DimPlot(SeuratObj, cols = MyPalette, group.by = "SCT_snn_res.0.3", 
        label = T, label.box = T, repel = T, label.color = "white", label.size = 4)
```


##    T lymphocytes

```{r ,fig.width=10, fig.height=7}
try(VlnPlot(SeuratObj, 
            features = c("CD3D","CD4","CD8A","IL7R"), 
            ncol = 2, 
            group.by = "SCT_snn_res.0.3",
            split.by = "hash.ID",
            cols = MyPalette))+
  theme(legend.position = "right",
        legend.justification = "top")

try(FeaturePlot(SeuratObj, features = c("CD3D","CD4","CD8A","IL7R"), cols = c("#EEEEEE","#221166")))
```


##    Other cell types

Basically here, we're looking for

-   Monocytes & macrophages
-   NK cells
-   B Lymphocytes
-   Myeloid cells & some others

```{r ,fig.width=12, fig.height=9}
try(VlnPlot(SeuratObj, 
            features = c("CD14","LYZ","NKG7","GNLY","MS4A1","MPO","EPCAM","CD34"), 
            ncol = 3, 
            group.by = "SCT_snn_res.0.3",
            split.by = "hash.ID",
            cols = MyPalette))+
  theme(legend.position = "right",
        legend.justification = "top")
```

As for the feature plots :

```{r ,fig.width=12, fig.height=9}
try(FeaturePlot(SeuratObj, 
                features = c("CD14","LYZ","NKG7","GNLY","MS4A1","MPO","EPCAM","CD34"),
                ncol = 3,
                cols = c("#EEEEEE","#221166")))
```






#   Clonotypes inspection   {.tabset .tabset-pills}

[Now we'll inspect the distribution of the clonotypes at the general condition (diag/rel) and among our clusters]{style='color:darkred ; font-size:120%'}

Before that, let's take a look at their distribution regardless of any clustering

```{r, rows.print=8}
SeuratObj@meta.data %>% 
  dplyr::count(clone_id) %>% 
  dplyr::arrange(desc(n)) %>% 
  head(15)
```


```{r}
SeuratObj@meta.data %>% 
  ggplot()+
  geom_point(aes(x= umap_1,
                 y= umap_2,
                 colour= clone_id),
             size = 0.7,
             alpha= 0.8)+
  My_ggplot_theme()+
  theme(legend.position = "none")
```


We have a lot of clonotypes marked in only 3 cells or less

[We'll also group them together, as for the unique clonotypes, they're considered healthy cells]{style='font-weight:600 ; color:navy'}

```{r ,rows.print=10}
unique_cl <- SeuratObj@meta.data %>% 
  dplyr::count(clone_id) %>% 
  dplyr::filter(n == 1) %>% 
  dplyr::pull(clone_id)

other_cl <- SeuratObj@meta.data %>% 
  dplyr::count(clone_id) %>% 
  dplyr::filter(n > 1 & n <= 3) %>% 
  dplyr::pull(clone_id) 

SeuratObj@meta.data <- SeuratObj@meta.data %>% 
  dplyr::mutate(clone_id = case_when(clone_id %in% unique_cl ~ "unique",
                                     clone_id %in% other_cl ~ "other",
                                     TRUE ~ clone_id))

SeuratObj@meta.data %>% 
  dplyr::count(clone_id) %>% 
  dplyr::arrange(desc(n)) %>% 
  head(10)
```

```{r}
SeuratObj@meta.data %>% 
  ggplot()+
  geom_point(aes(x= umap_1,
                 y= umap_2,
                 colour= clone_id),
             size = 0.7,
             alpha= 0.8)+
  My_ggplot_theme()+
  scale_colour_manual(values = MyPalette)+
  guides(colour = guide_legend(override.aes = list(size = 2.5)))
```



[Now let's see the distribution among our conditions]{style='color:darkred ; font-size:120%'}

## General conditions

```{r}
DimPlot(SeuratObj, cols = MyPalette, group.by = c("hash.ID"), 
        label = T, label.box = T, repel = T, label.color = "white", label.size = 4)
```

```{r ,rows.print=6}
SeuratObj@meta.data %>% 
  group_by(clone_id, hash.ID) %>% 
  summarise(count = n()) %>% 
  arrange(clone_id) %>%
  head(15)
```

```{r}
SeuratObj@meta.data %>% 
  group_by(clone_id, hash.ID) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) %>% 
  head(15) %>% 
  ggplot()+
  geom_bar(aes(x= hash.ID,
               y= count,
               fill= clone_id),
           stat = "identity",
           width = 0.9,
           position = "stack")+
  scale_fill_manual(values = MyPalette)+
  labs(title = "Distribution of the top clonotypes",
       x = "",
       y = "Count")+
  theme_bw()+
  theme(legend.title = element_text(size = 12, face = "bold", colour = "darkred"),
        legend.justification = "top",
        legend.text = element_text(colour = "navy"),
        axis.title = element_text(hjust = 0, size = 12, color = "darkred", face = "bold"),
        axis.text = element_text(size = 11))
```



##    Clusters

```{r}
DimPlot(SeuratObj, cols = MyPalette, group.by = c("Clust_res0.3"), 
        label = T, label.box = T, repel = T, label.color = "white", label.size = 4)
```


```{r ,rows.print=8}
SeuratObj@meta.data %>% 
  dplyr::count(Clust_res0.3) %>% 
  arrange(desc(n))
```

And with the clonotypes :

```{r ,rows.print=10}
SeuratObj@meta.data %>% 
  group_by(clone_id, Clust_res0.3) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) -> df ; df
```

```{r ,fig.width=10}
df %>% 
  head(30) %>% 
  ggplot()+
  geom_bar(aes(x= Clust_res0.3,
               y= count,
               fill= clone_id),
           stat = "identity",
           width = 0.7,
           position = "stack")+
  scale_fill_manual(values = MyPalette)+
  scale_y_continuous(breaks = c(30,500,1000,1500))+
  labs(title = "Distribution of the top clonotypes",
       x = "",
       y = "Count")+
  theme_bw()+
  theme(legend.title = element_text(size = 12, face = "bold", colour = "darkred"),
        legend.justification = "top",
        legend.text = element_text(colour = "navy"),
        axis.title = element_text(hjust = 0, size = 12, color = "darkred", face = "bold"),
        axis.text.x = element_text(size = 11, angle = 0),
        axis.text.y = element_text(size = 11),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_line(colour = "darkred", linetype = "dashed"))

df %>% 
  head(30) %>% 
  ggplot()+
  geom_bar(aes(x= Clust_res0.3,
               y= count,
               fill= clone_id),
           stat = "identity",
           width = 0.7,
           position = "fill")+
  scale_fill_manual(values = MyPalette)+
  labs(title = "Distribution of the top clonotypes",
       x = "",
       y = "Proportion")+
  theme_bw()+
  theme(legend.title = element_text(size = 12, face = "bold", colour = "darkred"),
        legend.justification = "top",
        legend.text = element_text(colour = "navy"),
        axis.title = element_text(hjust = 0, size = 12, color = "darkred", face = "bold"),
        axis.text.x = element_text(size = 11, angle = 0),
        axis.text.y = element_text(size = 11))
```



#   Filtering

[Now let's filter unwanted cells]{style='color:darkred ; font-size:120%'}

Based on the observations from the cells and clonotype analysis :

-   __Cluster7/2R/1D__ : not enough cells 
-   __Cluster3/4/6__ : different cell types
-   also remove __unique clonotypes__ and the others (__with less than 3 cells__)

```{r}
cells_rm <- WhichCells(SeuratObj, 
                       expression = umap_1 > 5 | umap_2 > 6 | 
                         Clust_res0.3 %in% c("1D","2R") | clone_id %in% c("unique","other"))
SeuratObj_filt <- subset(SeuratObj, cells = setdiff(Cells(SeuratObj),cells_rm))
```

```{r ,fig.width=12}
DimPlot(SeuratObj_filt, group.by = c("hash.ID","Clust_res0.3"), cols = MyPalette,
        label = T, label.box = T, repel = T, label.color = "white", label.size = 5)
```

[We'll personalize the clusters annotation according to the cell conditions with each cluster]{style='color:navy ; font-size:120% ; font-weight:600'} 

```{r}
SeuratObj_filt@meta.data <- SeuratObj_filt@meta.data %>% 
  mutate(MyClusters = case_when(Clust_res0.3 %in% c("2D") ~ "Diag1",
                                Clust_res0.3 %in% c("0R","1R") & umap_1 > 0.2 & umap_2 > -2.5 & umap_2 < 0 ~ "Rel_in_Diag1",
                                Clust_res0.3 %in% c("1R","0R") ~ "Rel1",
                                Clust_res0.3 %in% c("0D") & umap_1 <= 0 ~ "Diag_in_Rel1",
                                Clust_res0.3 == "5R" ~ "Rel2",
                                Clust_res0.3 == "5D" ~ "Diag2",
                                TRUE ~ "others"))

DimPlot(SeuratObj_filt, cols = MyPalette, group.by = c("MyClusters"), 
        label = T, label.box = T, repel = T, label.color = "white", label.size = 4)
```

```{r}
SeuratObj_filt@meta.data %>% 
  dplyr::count(MyClusters) %>% 
  dplyr::arrange(desc(n))
```


And now, let's inspect the distribution of the clonotypes again

```{r ,rows.print=8}
SeuratObj_filt@meta.data %>% 
  group_by(clone_id, MyClusters) %>% 
  summarise(count = n()) %>% 
  arrange(MyClusters) -> df ; df
```

Next, we'll do a final filtering removing all small groups with not enough cells

```{r}
cells_rm <- WhichCells(SeuratObj_filt, 
                       expression = MyClusters %in% c("Diag_in_Rel1","Rel_in_Diag1","others") | 
                         umap_1 > 0 & umap_2 > 2)
SeuratObj_filt <- subset(SeuratObj_filt, cells = setdiff(Cells(SeuratObj_filt),cells_rm))
```

```{r}
DimPlot(SeuratObj_filt, cols = MyPalette, group.by = c("MyClusters"), 
        label = T, label.box = T, repel = T, label.color = "white", label.size = 4)
```


And finally here's the summary of our top clonotypes remaining !

```{r ,rows.print=6}
SeuratObj_filt@meta.data %>% 
  group_by(clone_id, MyClusters) %>% 
  summarise(count = n()) %>% 
  arrange(clone_id) -> df ; df
```

As we can see __clonotype4 & 3__ are still insignificant in this library

```{r}
df %>% 
  ggplot()+
  geom_bar(aes(x= MyClusters,
               y= count,
               fill= clone_id),
           stat = "identity",
           width = 0.7,
           position = "stack")+
  scale_fill_manual(values = MyPalette)+
  labs(title = "Distribution of the top clonotypes",
       x = "",
       y = "Count")+
  theme_bw()+
  theme(legend.title = element_text(size = 12, face = "bold", colour = "darkred"),
        legend.justification = "top",
        legend.text = element_text(colour = "navy"),
        axis.title = element_text(hjust = 0, size = 12, color = "darkred", face = "bold"),
        axis.text.x = element_text(size = 11, angle = 0),
        axis.text.y = element_text(size = 11),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_line(colour = "darkred", linetype = "dashed"))
  
```

```{r}
SeuratObj_filt@meta.data %>% 
  mutate(TopClones = case_when(
    is.na(clone_id) ~ "Missing",
    clone_id %in% c("clonotype3","clonotype4") ~ "cl3-4",
    TRUE ~ clone_id
  )) -> SeuratObj_filt@meta.data
```


#   TCR inspections   {.tabset .tabset-pills}

[Now, let's inspect the VDJ expression :]

-   Among our conditions 
-   Among the clonotypes


##    Among the conditions

```{r ,fig.width=8}
DimPlot(SeuratObj_filt,
        group.by = "v_call",
        split.by = "MyClusters",
        cols = MyPalette)
```

Below is the summary of the remaining VDJ (V)

```{r ,rows.print=10}
SeuratObj_filt@meta.data %>% 
  dplyr::group_by(MyClusters, v_call) %>% 
  dplyr::summarise(count = n()) %>% 
  dplyr::arrange(desc(count)) -> df ; df
```

```{r}
df %>% 
  ggplot()+
  geom_bar(aes(x= MyClusters,
               y= count,
               fill= v_call),
           stat = "identity",
           width = 0.7,
           position = "stack")+
  scale_fill_manual(values = MyPalette)+
  labs(title = "VDJ distribution",
       x = "",
       y = "Count")+
  theme_bw()+
  theme(legend.title = element_text(size = 12, face = "bold", colour = "darkred"),
        legend.justification = "top",
        legend.text = element_text(colour = "navy"),
        axis.title = element_text(hjust = 0, size = 12, color = "darkred", face = "bold"),
        axis.text.x = element_text(size = 11, angle = 0),
        axis.text.y = element_text(size = 11),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_line(colour = "darkred", linetype = "dashed"))
```


##    Among the clonotypes

```{r ,fig.width=12}
DimPlot(SeuratObj_filt,
        group.by = "v_call",
        split.by = "TopClones",
        cols = MyPalette)
```

```{r ,rows.print=6}
SeuratObj_filt@meta.data %>% 
  dplyr::group_by(TopClones, v_call) %>% 
  dplyr::summarise(count = n()) %>% 
  dplyr::arrange(desc(count)) -> df ; df
```

```{r}
df %>% 
  ggplot()+
  geom_bar(aes(x= TopClones,
               y= count,
               fill= v_call),
           stat = "identity",
           width = 0.7,
           position = "stack")+
  scale_fill_manual(values = MyPalette)+
  labs(title = "VDJ distribution",
       x = "",
       y = "Count")+
  theme_bw()+
  theme(legend.title = element_text(size = 12, face = "bold", colour = "darkred"),
        legend.justification = "top",
        legend.text = element_text(colour = "navy"),
        axis.title = element_text(hjust = 0, size = 12, color = "darkred", face = "bold"),
        axis.text.x = element_text(size = 11, angle = 0),
        axis.text.y = element_text(size = 11),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_line(colour = "darkred", linetype = "dashed"))
```

```{r}
SeuratObj_filt@meta.data %>% 
  mutate(TopFeatures = case_when(
    is.na(v_call) ~ "Missing",
    v_call %in% c("TRAV13-1","TRAV13-2") ~ "Other",
    TRUE ~ v_call
  )) -> SeuratObj_filt@meta.data
```

```{r}
SeuratObj_filt@meta.data$TopClones_Condition <- str_c(
  paste0(str_split_i(SeuratObj_filt$TopClones,"onotype",1),
         str_split_i(SeuratObj_filt$TopClones,"onotype",2)), 
  substr(SeuratObj_filt$hash.ID,1,1),
  sep = "_"
)

SeuratObj_filt@meta.data$TopClones_Cond_TCR <- str_c(
  paste0(str_split_i(SeuratObj_filt$TopClones,"onotype",1),
         str_split_i(SeuratObj_filt$TopClones,"onotype",2)), 
  substr(SeuratObj_filt$hash.ID,1,1),
  SeuratObj_filt$v_call,
  sep = "_"
)

SeuratObj_filt@meta.data$Clusters_TopClones <- str_c(
  SeuratObj_filt$MyClusters,
  paste0(str_split_i(SeuratObj_filt$TopClones,"onotype",1),
         str_split_i(SeuratObj_filt$TopClones,"onotype",2)),
  sep = "_"
)

SeuratObj_filt@meta.data$Clone_TCR_Clust <- sub(pattern = "MissingNA_Missing",
                                                replacement = "Missing",
                                                x = str_c(SeuratObj_filt$Clusters_TopClones,
                                                          SeuratObj_filt$TopFeatures,
                                                          sep = "_"))


```

#   Expression plots

Now let's inspect the expression of some features including the vdj

```{r ,fig.height=5, fig.width=9}
FeaturePlot(SeuratObj_filt,
            features = c("TRAV8-3"))+
  scale_colour_gradientn(colours = c("white","lightgray","cadetblue","darkred"),
                         values = c(0,0.2,0.6,1),
                         limits = c(0,4)) -> p1

FeaturePlot(SeuratObj_filt,
            features = c("TRBV24-1"))+
  scale_colour_gradientn(colours = c("white","lightgray","cadetblue","darkred"),
                         values = c(0,0.2,0.6,1),
                         limits = c(0,4)) -> p2

cowplot::plot_grid(p1,p2, ncol = 2, labels = letters[1:4], label_size = 16)
```


```{r ,fig.width=10}
FeaturePlot(SeuratObj_filt, features = c("TRAV8-3","TRBV24-1"), blend = T, cols = c("lightgray","red","purple"))
```

```{r}
SeuratObj_filt@meta.data <- SeuratObj_filt@meta.data %>% 
  mutate(`_R1R2_vs_D` = case_when(MyClusters %in% c("Diag1","Diag2") ~ "Diag",
                                TRUE ~ MyClusters),
         `_cl2R_TCR_vs_D` = case_when(TopClones_Cond_TCR == "cl2_R_TRAV8-3" ~ "cl2_R_TRAV8-3",
                                      TopClones_Cond_TCR == "cl2_R_TRBV24-1" ~ "cl2_R_TRBV24-1",
                                      TRUE ~ hash.ID),
         `_cl2R_TCR_vs_D1D2` = case_when(TopClones_Cond_TCR == "cl2_R_TRAV8-3" ~ "cl2_R_TRAV8-3",
                                      TopClones_Cond_TCR == "cl2_R_TRBV24-1" ~ "cl2_R_TRBV24-1",
                                      TRUE ~ MyClusters),
         `_cl2R_TCR_vs_cl2D` = case_when(TopClones_Cond_TCR == "cl2_R_TRAV8-3" ~ "cl2_R_TRAV8-3",
                                         TopClones_Cond_TCR == "cl2_R_TRBV24-1" ~ "cl2_R_TRBV24-1",
                                         TRUE ~ TopClones_Condition)
         )
```



```{r}
saveRDS(SeuratObj_filt, file = paste0(out_seurat_dir,"Seurat3.rds"))
```

```{r ,echo=TRUE}
devtools::session_info()
```








































































































